title: "Assigment"
author: "marinac"
date: "20 September 2015"
output: html_document


##Loading and preprocessing the data##

```{r, echo=TRUE}
fileUrl<-"http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl,destfile="/Users/marina_casarino/Desktop/coursera/data.zip")
data<-read.csv(unzip("data.zip"))
data$date <- as.Date(data$date)
```


##What is mean total number of steps taken per day?##

**Histogram of the total number of steps taken each day**

```{r, echo=TRUE}
library(dplyr)
daily_data<-group_by(data,date)
new_data<-summarize(daily_data, steps=sum(steps,na.rm=TRUE))
hist(new_data$steps,col="grey",main= "Histogram of Number Step per Day",xlab="Number of Steps")
```

**Mean and Median total number of steps taken per day**

```{r, echo=TRUE}
mean_steps<-mean(new_data$steps, na.rm=TRUE )
median_steps<-median(new_data$steps, na.rm=TRUE )
```

The mean of daily steps is: `r mean_steps` and median of daily steps is: `r median_steps`


##What is the average daily activity pattern?##

**Line plot of average daily pattern**

```{r, echo=TRUE}
interval_data<-group_by(data,interval)
new_data2<-summarize(interval_data, steps=mean(steps,na.rm=TRUE))
plot(new_data2,type="l", main="Averaged Number of Steps per Interval ")
```

**Maximum number of steps per interval**

```{r, echo=TRUE}
max_value<-filter(new_data2,steps==max(steps))
max_value_interval<-max_value$interval
max_value_steps<-max_value$steps
```

The maximun average value of steps is: `r max_value$steps` and ocurr in the interval: `r max_value$interval`


##Imputing missing values##

**Missing Values**

```{r, echo=TRUE}
number_empty_values<-sum(is.na(data$steps))
```

The number of missing values for steps id: `r number_empty_values`

**Creating a new-completed dataset, by replacing the empty steps values by their average value**

```{r, echo=TRUE}
#selecting the data#
data_complete<-filter(data,is.na(steps)==FALSE)
data_incomplete<-filter(data,is.na(steps)==TRUE)
#calculating the values to replace NA#
interval_complete_data<-group_by(data_complete,interval)
new_data3<-summarize(interval_data, steps=mean(steps,na.rm=TRUE))
merged_data<-merge (select(data_incomplete,-steps),new_data3,by="interval")
#creating new dataset by merging the original set of complete data with the set of replaced values#
new_dataset<-rbind(merged_data,data_complete)
```

**New Histogram with Completed Data**

```{r, echo=TRUE}
new_daily_data<-group_by(new_dataset,date)
new_completed_data<-summarize(new_daily_data, steps=sum(steps))
hist(new_completed_data$steps,col="light blue",main= "New Histogram of Number Step per Day",xlab="Number of Steps")
```

**Mean and Median total number of steps taken per day using completed data**

```{r, echo=TRUE}
new_mean_steps<-mean(new_completed_data$steps )
new_median_steps<-median(new_completed_data$steps)
```

The mean of daily steps is: `r new_mean_steps` and median of daily steps is: `r new_median_steps`


##Differences in activity patterns between weekdays and weekends##

**Creating new dataset**

```{r, echo=TRUE}
#adding weekdays column#
new_dataset<-rbind(merged_data,data_complete)
new_database_with_days<-mutate(new_dataset,day=weekdays(date))
#selecting the data#
new_database_weekend<-mutate(filter(new_database_with_days,day=="Saturday"|day=="Sunday"))
new_interval_database_weekend<-summarize(group_by(new_database_weekend,interval),steps=mean(steps),type_of_day="weekend")
new_database_weekday<-mutate(filter(new_database_with_days,day!="Saturday" & day!="Sunday"))
new_interval_database_weekday<-summarize(group_by(new_database_weekday,interval),steps=mean(steps),type_of_day="weekday")
#final database with interval/average steps per interval/type_of_day#
final_database<-rbind(new_interval_database_weekend,new_interval_database_weekday)
```

**Panel Plot**

```{r, echo=TRUE}
library(lattice)
xyplot(final_database$steps ~ final_database$interval | final_database$type_of_day, data = final_database, type="l",main="Differences in Activity Patterns between Weekdays and Weekends",xlab="Interval",ylab="Number of Steps",layout = c(1,2))
```
